### 2021-04-25 tex/paper/images/tikz/diagram.tex
Diagram appeared
First list of items was changed on the diagram. Affiliations were updated.

\documentclass[tikz,border=5]{standalone}

\usepackage{tikz}
\usetikzlibrary{arrows}

\begin{document}

\begin{tikzpicture}

\draw (0.0,0) node[rectangle, line width=1.15pt, fill=blue!20, minimum height=1cm, minimum width=2.25cm, draw] {Processing};

\draw (3.5,0) node[rectangle, line width=1.15pt, fill=blue!20, minimum height=1cm, minimum width=2.25cm, draw] {Vectorization};

\draw (7.0,0) node[rectangle, line width=1.15pt, fill=blue!20, minimum height=1cm, minimum width=2.25cm, draw] {Classification};

\draw[-latex', line width=1.15pt] (1.25,0) -- (2.25,0);

\draw[-latex', line width=1.15pt] (4.75,0) -- (5.75,0);

\end{tikzpicture}
\end{document}